---
title: "version_2"
author: "John Goldin"
date: "5/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(sparkline)
library(kableExtra)
```

The hope is that this will grow into a replacement blog post for Connecticut Covid-19 stats, but baased
on RMarkdown rather than simple markdown.


Assumes all the info in daily_ct_stats.R is available.
Using week_setup to get as-of dates,
age_town_acs to get populaton and town geometries
and dph_towns to get Covid-19 cases and deaths




```{r setup_town_info}
town_info_plus <- town_info %>% 
  left_join(dph_towns, by = "town") %>% 
  filter(date %in% week_setup$end_period)
```

## Including Plots

You can also embed plots, for example:

```{r t, echo=FALSE}
# note: need aes(geometry = geometry) because operation that turns sf object into tibble loses someting that specifies geometry column for ggplot
p <- ggplot() + 
  geom_sf(data = town_info, aes(geometry = geometry),
  fill = "gray", colour = "darkgray", show.legend = FALSE) +
  geom_sf(data = dph_pop, colour = "yellow", fill = NA) +
  # geom_sf_text(data = county_centroid, aes(label = county), color = "yellow") +
  coord_sf(datum = NA, label_axes = "----") +
  xlab("") + ylab("") + theme_minimal() +
  labs(title = "Connecticut Counties and Towns", 
       subtitle = "county boundary in yellow",
       caption = "Source: US Census, tidycensus package")  
```

```{r}
county_centroid <- st_centroid(dph_pop) # use to place town labels
county_map <- ggplot() +
  geom_sf(data = dph_pop, aes(fill = cases_per_pop)) +
  scale_fill_gradient(low = "white", high = "grey") +
  geom_sf_text(data = county_centroid, aes(label = paste0(county, "\n", per_pop_label, " per 100K")), color = "black", size = 3) +
  coord_sf(datum = NA, label_axes = "----") +
  xlab("") + ylab("") + theme_minimal()  +
  # theme(legend.position = "none") +
  labs(title = "Cumulative Confirmed Cases by Connecticut County",
       subtitle = "cases per 100K of population for most recent report date",
       fill = "cases per 100K",
       caption = "Source: US Census, tidycensus package")
```


```{r, eval = FALSE}
ct <- ct %>% mutate(wday = factor(weekdays(date),
                                  levels = c("Monday", "Tuesday",
                                             "Wednesday", "Thursday", "Friday", "Saturday","Sunday")))
p <- ggplot(data = ct %>% filter(county == "Fairfield"), aes(x = date, y = new_cases)) +
  geom_col(aes(fill = factor(wday))) + 
  scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.9, option = "viridis")+ 
  geom_line(aes(y = rnew_cases), size = 1) +
  xlab(NULL) + ylab("New Confirmed Cases Per Day")

p_deaths <- ggplot(data = ct %>% filter(county == "Fairfield"), aes(x = date, y = new_deaths)) +
  # geom_col(aes(fill = factor(wday))) +
  # scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.9, option = "viridis") +
  geom_line(colour = "lightgreen", size = 0.2) +
  labs(fill = "Day of Report") +
  geom_line(aes(y = rnew_deaths), size = 1, colour = "darkgreen") +
  xlab(NULL) + ylab("Covid19 Deaths Per Day") +
  labs(title = "Connecticut Covid-19 Deaths Per Day", 
       subtitle = "Trend line is Smoothed Over the Last Seven Days",
       caption = "Source: US Census, tidycensus package") +
  scale_x_date(limits = c(ymd("2020-03-14"), NA), minor_breaks = NULL,
               date_breaks = "1 month", date_labels = "%B") +
  # scale_y_log10() +
  theme_minimal() 

p_deaths_cases1 <- ggplot(data = ct %>% filter(county == "Fairfield"), aes(x = date, y = new_deaths)) +
  # geom_col(aes(fill = factor(wday))) + 
  # scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.9, option = "viridis") + 
  geom_line(colour = "lightgreen", size = 0.2) +
  labs(fill = "Day of Report") +
  geom_line(aes(y = rnew_deaths), size = 1, colour = "darkgreen") +
  xlab(NULL) + ylab("Covid19 Deaths Per Day") +
  labs(title = "Connecticut Covid-19 Deaths Per Day", 
       subtitle = "Trend line is Smoothed Over the Last Seven Days",
       caption = "Source: US Census, tidycensus package") +
  scale_x_date(limits = c(ymd("2020-03-14"), NA), minor_breaks = NULL,
               date_breaks = "1 month", date_labels = "%B") +
  geom_line(aes(y = new_cases), size = 0.2, colour = "lightblue") +
  geom_line(aes(y = rnew_cases), size = 1, colour = "darkblue") +
  scale_y_log10() +
  theme_minimal()

p_state_new <- dph_total %>% select(date, cases = rnew_cases, deaths = rnew_deaths)  %>%  pivot_longer(cols = c("cases", "deaths"), names_to = "type", values_to = "rolling") %>% 
  inner_join(
    dph_total %>% select(date, cases = new_cases, deaths = new_deaths)  %>%
      pivot_longer(cols = c("cases", "deaths"), 
                   names_to = "type", values_to = "new"), 
    by = c("date", "type")  ) %>% 
  mutate(type = factor(type, levels = c("cases", "deaths"), 
                       labels = c("New Confirmed Cases", "Covid-19 Deaths"))) %>% 
  ggplot(aes(x = date, colour = type)) +
  geom_line(aes(y = new), size = 0.2) +
  geom_line(aes(y = rolling), size = 1) +
  facet_wrap(~ type, scales = "free") +
  # scale_x_date(limits = c(ymd("2020-03-14"), NA), minor_breaks = NULL,
  #              date_breaks = "1 month", date_labels = "%B") +
  xlab(NULL) + ylab(NULL) +
  labs(title = "Connecticut Covid-19 New Cases and Deaths Per Day", 
       subtitle = "Trend line is Smoothed Over the Last Seven Days",
       caption = "Source: Connecticut Department of Public Health") +
  theme_minimal() +
  theme(legend.position = "none", strip.text = element_text(size = 11))

counties_new_data <- dph_counties %>% select(date, county, cases = rnew_cases, deaths = rnew_deaths)  %>%  
  pivot_longer(cols = c("cases", "deaths"), names_to = "type", values_to = "rolling") %>% 
  inner_join(
    dph_counties %>% select(date, county, cases = new_cases, deaths = new_deaths)  %>%
      pivot_longer(cols = c("cases", "deaths"), names_to = "type", values_to = "new"), 
    by = c("county", "date", "type")  ) %>% 
  mutate(type = factor(type, levels = c("cases", "deaths"), 
                       labels = c("New Confirmed Cases", "Covid-19 Deaths"))) 
aplot1 <- function(d, pick_counties) {
  d %>% filter(county %in% pick_counties) %>% 
  ggplot(aes(x = date, colour = type)) +
  geom_line(aes(y = new), size = 0.2) +
  geom_line(aes(y = rolling), size = 1) +
  facet_wrap(county ~ type, scales = "free",nrow = 4) +
  # scale_x_date(limits = c(ymd("2020-03-14"), NA), minor_breaks = NULL,
  #              date_breaks = "1 month", date_labels = "%B") +
  xlab(NULL) + ylab(NULL) +
  labs(title = "Connecticut Covid-19 New Cases and Deaths Per Day", 
       subtitle = "Trend line is Smoothed Over the Last Seven Days",
       caption = "Source: Connecticut Department of Public Health") +
  theme_minimal() +
  theme(legend.position = "none", strip.text = element_text(size = 11))
}
p_counties_new1 <- aplot1(p_counties_new_data, pick_counties = c("New Haven", "Fairfield", "Hartford", "Litchfield"))
p_counties_new2 <- aplot1(p_counties_new_data, pick_counties = c("Tolland", "New London", "Middlesex", "Windham"))


p_counties_new2 <- counties_new_data %>%  
  ggplot(aes(x = date, colour = county)) +
  geom_line(aes(y = new), size = 0.2) +
  geom_line(aes(y = rolling), size = 1) +
  facet_wrap(~ type, scales = "free",nrow = 4) +
  # scale_x_date(limits = c(ymd("2020-03-14"), NA), minor_breaks = NULL,
  #              date_breaks = "1 month", date_labels = "%B") +
  xlab(NULL) + ylab(NULL) +
  labs(title = "Connecticut Covid-19 New Cases and Deaths Per Day", 
       subtitle = "Trend line is Smoothed Over the Last Seven Days",
       caption = "Source: Connecticut Department of Public Health") +
  theme_minimal() +
  theme(legend.position = "none", strip.text = element_text(size = 11))

p_counties_current <- ct %>% 
  ggplot(aes(x = date, colour = county)) +
  geom_line(aes(y = cases_per_100k), size = 0.8) +
  # scale_x_date(limits = c(ymd("2020-03-14"), NA), minor_breaks = NULL,
  #              date_breaks = "1 month", date_labels = "%B") +
  xlab(NULL) + ylab(NULL) +
  labs(title = "Connecticut Covid-19 Active Confirmed Cases per 100K of Population", 
       subtitle = "(Assumes cases remain active for two weeks.)",
       caption = "Source: Connecticut Department of Public Health") +
  theme_minimal() 
```



```{r sparkline_experiment}
test <- town_info_plus %>% 
  filter(town %in% c("Guilford", "New Haven", "Madison"),
         !is.na(rnew_cases))
sparkline(test$rnew_cases[3])
# sparkline(test$rnew_cases[2])
# sparkline(test$rnew_cases[3])
```


```{r}

```


