---
title: "Exploring Covid-19 Trends by Town"
author: "John Goldin"
date: "6/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(sparkline)
library(kableExtra)
library(formattable)
```
```{r setup_town_info, echo = FALSE}
path_to_post <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/content/post/2020-03-29-covid19-cases-in-connecticut/"
path_to_ctcorona <- "~/Documents/R_local_repos/ctcorona/"
load(paste0(path_to_ctcorona, "dph_datasets.RData"))
load(paste0(path_to_ctcorona, "census_population.RData"))
```

This table summarizes trends in towns with a population greater than 14,000. It's sorted by county and by category of town as
developed in the report [The Five Connecticuts](https://ctdatahaven.org/sites/ctdatahaven/files/UConnCPR%20Changing%20Demographics-5%20CTs%202004.pdf).

For each town it shows the population, the largest average number of daily new cases
(based on rolling 14 day average), and the percentage of deaths
attributed to nursing home residents. 

Next there are a series of [sparklines](https://en.wikipedia.org/wiki/Sparkline)
that show the trend in daily new cases and daily deaths (based on 
14 day moving average) and the same for the county containing that
town. 

The goal is to try to quickly scan the towns looking for patterns.


```{r, echo = FALSE, results = "asis"}
# THE ORIGINAL VERSION
# based on https://stackoverflow.com/questions/54244718/combining-sparkline-in-kableextra-table
# jQuery sparklines https://omnipotent.net/jquery.sparkline/#s-docs
county_sparklines <- dph_counties %>% 
  group_by(county) %>% 
  summarise(`county cases` = spk_chr(rnew_cases),
            `county deaths` = spk_chr(rnew_deaths), .groups = "drop")
test <- dph_towns %>%
  filter(total_pop > 14000,
         !is.na(rnew_cases))# %>%
test <- test %>% 
  # left_join(town_info %>% select(county, category, town)) %>% 
  group_by(county, category, town) %>% 
  # summarise(`max cases` = round(max(rnew_cases, na.rm = TRUE)),
  summarise(`max cases` = max(rnew_cases) %>% round(1),
    `town cases` = spk_chr(rnew_cases),
    `town deaths` = spk_chr(rnew_deaths), .groups = "drop") %>% 
  left_join(county_sparklines, by = "county") %>% 
  left_join(town_info %>% select(town, total_pop), by = "town") %>% 
  left_join(town_with_nursing %>% 
              filter(date == max(date)) %>% 
              select(town, nh_death_pct), by = "town") %>% 
  select(town, category, county, total_pop, nh_death_pct, `max cases`, `town cases`,
         `town deaths`, `county cases`, `county deaths`) %>% 
  formattable::format_table(
    # x = .,
    # formatters = list(align=c("l")),
    format = "html",
    align = c("l", "l", "l", "l", "c", "c", "c", "c")
  ) %>%
  # kable_styling("striped", full_width = F) %>%
  # kable_styling(full_width = F) %>%
  # group_rows("group1", 1, 2) %>%
  # group_rows("group1", 1, 2) %>%
  # group_rows("group2", 3,3) %>%
  htmltools::HTML() %>%
  shiny::div() %>%
  sparkline::spk_add_deps()
test
```




