---
title: Tracking Covid-19 in Connecticut
author: John Goldin
date: '2020-06-12'
slug: []
categories:
  - COVID19
tags:
  - COVID19
type: post
subtitle: ''
image: ''
draft: yes
output:
  blogdown::html_page:
    toc: yes
    fig_width: 7
    out.width: "70%" 
    fig.align: "center"
    fig.asp: 0.618
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, out.width = "100%",
                      knitr.table.format = "html")
# knitr.table.format is for kable. Needed this to get blogdown table to format correctly.
```
```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse, quietly = TRUE)
library(knitr, quietly = TRUE)
library(scales, quietly = TRUE)
library(kableExtra, quietly = TRUE)
library(sparkline)
```
```{r `load_basic_data`, echo = FALSE, message = FALSE}
path_to_post <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/content/post/2020-03-29-covid19-cases-in-connecticut/"
path_to_static_for_this_post <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/static/post/2020-06-12-tracking-covid-19-in-connecticut"
path_to_ctcorona <- "~/Documents/R_local_repos/ctcorona/data/"
load(paste0(path_to_ctcorona, "dph_datasets.RData"))
load(paste0(path_to_ctcorona, "census_population.RData"))

# from https://www.ctdatahaven.org/sites/ctdatahaven/files/UConnCPR%20Changing%20Demographics-5%20CTs%202004.pdf
# The Changing Demographics of Connecticut — 1990 to 2000. by Center for Population Research  May. 31, 2004
Five_Connecticuts <- read_delim(paste0(path_to_ctcorona, "Five_Connecticuts.txt"),"\t", escape_double = FALSE, trim_ws = TRUE) %>%
  mutate(category = factor(category, levels = c("Urban Core", "Urban Periphery",
                           "Wealthy", "Suburban", "Rural")))
load(paste0(path_to_ctcorona, "dph_datasets.RData"))

```
The [post I created back in March](https://johngoldin.com/post/2020-03-29-covid19-cases-in-connecticut/) 
aimed to track the growth of Covid-19 as it hit Connecticut.
Back then (at the end of March) I didn't realize that 
the greater New York area (including Connecticut)
would be the center of the first wave in the US. At that time what I was reading
focused on the nature of exponential growth of an epidemic. We were urged to
"flatten the curve" and there was an emphasis on the shape of the growth of
cumulative cases and deaths caused by the virus. I have been upading the figures
in that post each evening. During tht time the [Connecticut data portal](https://johngoldin.com/post/2020-03-29-covid19-cases-in-connecticut/)
has expanded, and I've built [some R code](https://github.com/johngoldin/ctcorona/blob/master/R/daily_ct_stats.R) 
that let's me easily update
that post. I'll use the same data infrastructure for this post as well.

### Tracking the Virus in Connecticut

This post focuses on what is happening recently. The aim is to
use the recent trend to get some idea of the near future. The first
plot will show the average number of new cases reported each day along
with a line that displays the rolling seven-day average. Typically there 
are day-of-the-week effects in the reporting so it's best to focus
on the seven-day average.

The plot show the history of new cases and also shows the 
actions in some of Governor Lamont's executive orders. 

The number of new cases peaked in mid April and then began to
decrease almost as rapidly. See [this site](https://rt.live/) for
an estimate by state of the effective value for 
R<sub>t</sub>, the average number of people 
who become infected by an infectious person. The site
is described in [this article](https://www.vox.com/recode/2020/4/21/21227855/coronavirus-spreading-by-state-instagram-effective-reproduction-rate) at Vox.com.

Note that as of June (and a month
previous), Connecticut, New York, and New Jersey have among the lowest
values for R<sub>t</sub> in the country. This is good news for us
in Connecticut.

```{r 'new_cases_plot', echo = FALSE, eval = TRUE}
## This imaged will be saved in the /post/*_files/ directory
## Use echo = FALSE if you want to hide the code for making the plot
exec_orders <- tibble(
  date = as.Date(c("2020-03-10", "2020-03-17", "2020-03-23", "2020-03-26", "2020-04-11", "2020-04-20", "2020-05-20", "2020-06-01", "2020-06-17")),
  label = c("prohibit large\ngatherings", "cancel\nclasses", "restrict\nbusiness", "5 or\nfewer",
            "covid-19\ndeath\nchanged", "masks\nequired",
            "phase 1\nreopen", "hair\nreopen", "phase 2\nreopen"),
  fudge = c(1, 0.4, 0.7, 0.9, 0.8, 0.9, 0.8, 0.8, 1)
) %>% 
  left_join(dph_total %>% select(date, rnew_cases, rnew_deaths, cases, deaths))

# try a version with understated bars
p_new_cases <- ggplot(data = dph_total %>% 
              # filter(county == "Fairfield") %>% 
              mutate(wday  = factor(weekdays(date),
                                  levels = c("Monday", "Tuesday",
                                             "Wednesday", "Thursday", "Friday",
                                             "Saturday","Sunday"))), 
            aes(x = date, y = new_cases)) +
  geom_col(fill = "gray95", colour = "gray90") + 
  # scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.7, option = "viridis") + 
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", date_labels = "%b %d") +
  geom_line(aes(y = rnew_cases), size = 1) +
  theme_minimal() + theme(legend.position = "none") +
  xlab(NULL) + ylab("New Cases Per Day") +
  geom_text(data = exec_orders %>% filter(date <= max(dph_total$date, na.rm = TRUE)), 
            aes(y = (rnew_cases + 500) * fudge , label = label),
            hjust = 0.5, vjust = 0, size = 3, colour = "black", nudge_y = 0) +
  ggtitle("Connecticut New Covid-19 Cases (with 7-day rolling average)")

p_daily_deaths <- ggplot(data = dph_total %>% 
              # filter(county == "Fairfield") %>% 
              mutate(wday  = factor(weekdays(date),
                                  levels = c("Monday", "Tuesday",
                                             "Wednesday", "Thursday", "Friday",
                                             "Saturday","Sunday"))), 
            aes(x = date, y = new_deaths)) +
  geom_col(fill = "gray95", colour = "gray90") + 
  # scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.7, option = "viridis") + 
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", date_labels = "%b %d") +
  geom_line(aes(y = rnew_deaths), size = 1) +
  theme_minimal() + theme(legend.position = "none") +
  xlab(NULL) + ylab("Deaths Per Day") +
  geom_text(data = exec_orders %>% filter(date <= max(dph_total$date, na.rm = TRUE)), 
            aes(y = (rnew_deaths + 75) * fudge , label = label),
            hjust = 0.5, vjust = 0, size = 3, colour = "black", nudge_y = 0) +
  ggtitle("Connecticut Covid-19 Daily Deaths (with 7-day rolling average)")
# load effective R from https://rt.live/
rt_live <- read_csv("https://d14wlfuexuxgcm.cloudfront.net/covid/rt.csv")
p_rt_live <- ggplot(data = rt_live %>% filter(region == "CT"), aes(x = date, y = mean)) +
  geom_ribbon(aes(ymin = lower_90, ymax = upper_90), fill = "grey95") +
  geom_ribbon(aes(ymin = lower_50, ymax = upper_50), fill = "grey80") +
  geom_line() + theme_minimal() + 
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(0, 1.5), breaks = c(0, 1), 
                     minor_breaks = seq(0.2, 1.4, 0.2)) +
  geom_hline(yintercept = 1, colour = "gray") +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", 
               date_labels = "%b %d", 
               limits = c(min(dph_total$date), max(dph_total$date))) +
  xlab(NULL) + ylab("effective R") +
  ggtitle("Effective R (as computed by https://rt.live/)")
```
```{r hospitalizations, echo = FALSE, out.width = "70%", fig.align = "center"}
p_hospitalizations <- ggplot(data = dph_total %>% filter(!is.na(hospital)),
                    aes(x = date, y = hospital)) +
  geom_line() + xlab(NULL) + # geom_point() + 
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", 
               date_labels = "%b %d", 
               limits = c(min(dph_total$date), NA)) +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("in hospital") +
  labs(
    title = "COVID19 Patients Currently in Hospital",
    subtitle = NULL,
    caption = "Source: Connecticut Hospital Association, via CT DPH"
  )
```
```{r p_new_cases, echo = FALSE}
print(p_new_cases)

```
```{r p_rt_live, echo = FALSE, fig.asp = 0.2}
print(p_rt_live)
```
```{r p_daily_deaths, echo = FALSE}
print(p_daily_deaths)

```
```{r p_hospitalizations, echo = FALSE}
print(p_hospitalizations)
```

These charts shows the counts of cases and deaths as they are reported by
the Connecticut Department of Public Health. Reports sometimes end up in
batches. There is a day of the week effect, with fewer cases on weekends.
Some of the unusually large peaks and valleys in these charts are due
to reporting process. The DPH has also begun producing reports based
on the date a sample was taken for a test and for the date of
death. That's a more accurate way to look at the change over time, but it
means the data for the most recent days are hard to interpret because
some data may be "in process" and not yet reported. In these charts I 
have used "date of report" rather than "date of sample taken" or "date of death"
because that gives me all of the recent data that is available
and because that is what most other data projects (such as the Covid-19 Project
or the New York Times) have been using.



