---
title: Tracking Covid-19 in Connecticut
author: John Goldin
date: '2020-06-19'
slug: []
categories:
  - COVID19
tags:
  - COVID19
type: post
subtitle: ''
image: ''
draft: yes
output:
#  blogdown::html_page:
  bookdown::html_document2:
    toc: yes
    number_sections: false
    # figure parameters based on recommendation in Hadley's book
    fig_width: 7
    out.width: "70%" 
    fig.align: "center"
    fig.asp: 0.618
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, out.width = "100%",
                      knitr.table.format = "html")
# knitr.table.format is for kable. Needed this to get blogdown table to format correctly.
```
```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse, quietly = TRUE)
library(knitr, quietly = TRUE)
library(scales, quietly = TRUE)
library(kableExtra, quietly = TRUE)
library(sparkline)
library(geofacet) # https://hafen.github.io/geofacet/
library(sf, quietly = TRUE)
options(tigris_use_cache = TRUE)
library(tigris, quietly = TRUE)
library(tidycensus, quietly = TRUE)
library(viridis, quietly = TRUE)
library(ggrepel)
options(tigris_class = "sf")
library(lubridate, quietly = TRUE)
library(janitor)
library(fuzzyjoin) # for interval_left_join
library(broom)
library(RSocrata)
library(RcppRoll)
```
```{r `load_basic_data`, echo = FALSE, message = FALSE}
path_to_post <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/content/post/2020-03-29-covid19-cases-in-connecticut/"
path_to_static_for_this_post <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/static/post/2020-06-19-tracking-covid-19-in-connecticut"
path_to_ctcorona <- "~/Documents/R_local_repos/ctcorona/data/"
path_to_ctcorona_r <- "~/Documents/R_local_repos/ctcorona/R/"
load(paste0(path_to_ctcorona, "dph_datasets.RData"))
load(paste0(path_to_ctcorona, "census_population.RData"))
source(paste0(path_to_ctcorona_r, "combine_dph_towns.R"))

# from https://www.ctdatahaven.org/sites/ctdatahaven/files/UConnCPR%20Changing%20Demographics-5%20CTs%202004.pdf
# The Changing Demographics of Connecticut — 1990 to 2000. by Center for Population Research  May. 31, 2004
Five_Connecticuts <- read_delim(paste0(path_to_ctcorona, "Five_Connecticuts.txt"),"\t", escape_double = FALSE, trim_ws = TRUE) %>%
  mutate(category = factor(category, levels = c("Urban Core", "Urban Periphery",
                           "Wealthy", "Suburban", "Rural")))
# note that prisons have dominant effect on Somers and Montville
load(paste0(path_to_ctcorona, "dph_datasets.RData"))
# load effective R from https://rt.live/
rt_live <- read_csv("https://d14wlfuexuxgcm.cloudfront.net/covid/rt.csv") %>% 
  mutate(state = case_when(
    region == "CT" ~ "Connecticut",
    region == "AZ" ~ "Arizona"
  )) %>% 
  group_by(region) %>% arrange(date) %>% 
  mutate(above_one = ifelse(last(mean) > 1, "< 1", ">= 1") %>% factor(levels = c(">= 1", "< 1"))) %>% 
  ungroup()
```

## Latest Connecticut Data from  the Department of Public Health
```{r box-score, echo = FALSE}
recent_date <- max(dph_total$date, na.rm = TRUE)
week_ago <- recent_date - 7
if (!(week_ago %in% dph_total$date)) week_ago <- NA
yesterday <- recent_date - 1
# if (!(yesterday %in% dph_total$date)) yesterday <- NA
# if (is.na(yesterday) || (length(dph_total$deaths[dph_total$date == yesterday]))) {
#   yesterday_deaths <- ""
# } else {yesterday_deaths <- dph_total$deaths[dph_total$date == recent_date] - dph_total$deaths[dph_total$date == yesterday]}
today_cases <- dph_total %>% filter(date == recent_date) %>% 
  select(cases) %>% pluck(1, .default = NA)
yesterday_cases <- today_cases - dph_total %>% filter(date == yesterday) %>% 
  select(cases) %>% pluck(1, .default = NA)
week_ago_cases <-  today_cases - dph_total %>% filter(date == week_ago) %>% 
  select(cases) %>% pluck(1, .default = NA)
today_deaths <- dph_total %>% filter(date == recent_date) %>% 
  select(deaths) %>% pluck(1, .default = NA)
yesterday_deaths <- today_deaths - dph_total %>% filter(date == yesterday) %>% 
  select(deaths) %>% pluck(1, .default = NA)
week_ago_deaths <-  today_deaths - dph_total %>% filter(date == week_ago) %>% 
  select(deaths) %>% pluck(1, .default = NA)
today_hospital <- dph_total %>% filter(date == recent_date) %>% 
  select(hospital) %>% pluck(1, .default = NA)
yesterday_hospital <- today_hospital - dph_total %>% filter(date == yesterday) %>% 
  select(hospital) %>% pluck(1, .default = NA)
week_ago_hospital <-  today_hospital - dph_total %>% filter(date == week_ago) %>% 
  select(hospital) %>% pluck(1, .default = NA)
today_tests <- dph_total %>% filter(date == recent_date) %>% select(covid_19_pcr_tests_reported) %>% pluck(1, .default = NA)
yesterday_tests <- today_tests - dph_total %>% filter(date == yesterday) %>% select(covid_19_pcr_tests_reported) %>% pluck(1, .default = NA)
week_ago_tests <-  today_tests - dph_total %>% filter(date == week_ago) %>% select(covid_19_pcr_tests_reported) %>% pluck(1, .default = NA)
test_rate <- ((dph_total %>% filter(date == recent_date) %>% 
                 select(confirmedcases) %>% pluck(1, .default = NA)) -
                (dph_total %>% filter(date == yesterday) %>% 
                   select(confirmedcases) %>% pluck(1, .default = NA))) / 
  ((dph_total %>% filter(date == recent_date) %>% 
      select(covid_19_pcr_tests_reported) %>% pluck(1, .default = NA)) -
     (dph_total %>% filter(date == yesterday) %>% 
        select(covid_19_pcr_tests_reported) %>% pluck(1, .default = NA)))
recent_nursing_date <- max(dph_nursing_cases$date, na.rm = TRUE)
nursing_previous_date <- dph_nursing_cases %>% select(date) %>% arrange(desc(date)) %>% unique()
nursing_previous_date <- nursing_previous_date$date[2]
nursing_hm_cases <- dph_nursing_cases %>% filter(date == recent_nursing_date) %>% 
  select(nh_cases) %>% pluck(1, .default = NA) %>% sum(na.rm = TRUE)
nh_previous_cases <-  nursing_hm_cases - dph_nursing_cases %>% filter(date == nursing_previous_date) %>% 
  select(nh_cases) %>% pluck(1, .default = NA) %>% sum(na.rm = TRUE)
nursing_hm_deaths <- dph_nursing_cases %>% filter(date == recent_nursing_date) %>% 
  select(nh_deaths) %>% pluck(1, .default = NA) %>% sum(na.rm = TRUE)
nh_previous_deaths <-  nursing_hm_deaths - dph_nursing_cases %>% filter(date == nursing_previous_date) %>% 
  select(nh_deaths) %>% pluck(1, .default = NA) %>% sum(na.rm = TRUE)
# use: sprintf("%+3d %%", x)
box_fmt <- function(x) {
  case_when(
    (is.na(x) | is.null(x) | (length(x) == 0)) ~ " ",
    x < 0 ~ comma(x),
    x > 0 ~ comma(x, prefix = "+"),
    TRUE ~  "0"
  )
}

###########################################################################
# Box score comes here:
box_score <- tribble(
  ~` `, ~`today`, ~`yesterday`, ~`week_ago`,
  "Cases", today_cases, yesterday_cases, week_ago_cases,
  "Deaths", today_deaths, yesterday_deaths, week_ago_deaths,
  "Hospitalizations", today_hospital, yesterday_hospital, week_ago_hospital,
  "Tests",    today_tests, yesterday_tests, week_ago_tests,
  "--- ", NA , NA, NA,
  "Nursing home cases", nursing_hm_cases, NA, nh_previous_cases,
  "Nursing home deaths", nursing_hm_deaths, NA, nh_previous_deaths
) %>% 
    mutate(yesterday = map_chr(yesterday, box_fmt),
           week_ago = map_chr(week_ago, box_fmt), today = ifelse(is.na(today), "", comma(today)))

  # mutate(across(c(yesterday, week_ago), box_fmt), today = ifelse(is.na(today), "", comma(today)))

kable(box_score, 
      col.names = c(" ", "Total Today", "Change from\nYesterday", "Change from\nWeek Ago"),
      format.args = list(big.mark = ","),
      align = "lrrr") %>% 
  kableExtra::column_spec(2:4, width_max = "1.5in") %>% 
  kableExtra::column_spec(1, bold = TRUE, width_min = "1.5in")
```
 **Rate of positive tests:    **  `r round(test_rate * 100, 1.1)`%
 
 PDF report from the Connecticut Deprtment of Public Health:[`r dph_reports$name[1]`](`r dph_reports$url[1]`)
 
 Source: Multiple tables from [Connecticut Open Data](https://data.ct.gov/stories/s/COVID-19-data/wa3g-tfvc/)  
 -------
 
A [post I created back in March](https://johngoldin.com/post/2020-03-29-covid19-cases-in-connecticut/) 
aimed to track the growth of Covid-19 as it hit Connecticut.
Back then at the end of March I did not realize that 
the greater New York area (including Connecticut)
would be the center of the first wave in the US. At that time what I was reading
focused on the nature of exponential growth of an epidemic. We were urged to
"flatten the curve" and that shaped the way I presented the
data in that post. I have been updating the figures
in that post each evening. During that time the [Connecticut data portal](https://data.ct.gov/stories/s/COVID-19-data/wa3g-tfvc/)
has expanded considerably, and I've built [some R code](https://github.com/johngoldin/ctcorona/blob/master/R/daily_ct_stats.R) 
that let's me easily update
that post. I'll use the same data infrastructure for this post as well.
This post is much more oriented toward tracking the current status of 
the epidemic in Connecticut.

```{r 'new-cases-plot', echo = FALSE, eval = TRUE}
## This imaged will be saved in the /post/*_files/ directory
## Use echo = FALSE if you want to hide the code for making the plot
exec_orders <- tibble(
  date = as.Date(c("2020-03-10", "2020-03-17", "2020-03-23", "2020-03-26", "2020-04-11", "2020-04-20", "2020-05-20", "2020-06-01", "2020-06-17")),
  label = c("prohibit large\ngatherings", "cancel\nclasses", "restrict\nbusiness", "5 or\nfewer",
            "covid-19\ndeath\nchanged", "masks\nrequired",
            "phase 1\nreopen", "hair\nreopen", "phase 2\nreopen"),
  fudge = c(1, 0.4, 0.7, 0.9, 0.8, 0.9, 0.8, 0.8, 1)
) %>% 
  left_join(dph_total %>% select(date, rnew_cases, rnew_deaths, cases, deaths))

# try a version with understated bars
p_new_cases <- ggplot(data = dph_total %>% 
              # filter(county == "Fairfield") %>% 
              mutate(wday  = factor(weekdays(date),
                                  levels = c("Monday", "Tuesday",
                                             "Wednesday", "Thursday", "Friday",
                                             "Saturday","Sunday"))), 
            aes(x = date, y = new_cases)) +
  geom_col(fill = "gray95", colour = "gray90") + 
  # scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.7, option = "viridis") + 
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", date_labels = "%b %d") +
  geom_line(aes(y = rnew_cases), size = 1) +
  theme_minimal() + theme(legend.position = "none") +
  xlab(NULL) + ylab("New Cases Per Day") +
  geom_text(data = exec_orders %>% filter(date <= max(dph_total$date, na.rm = TRUE)), 
            aes(y = (rnew_cases + 500) * fudge , label = label),
            hjust = 0.5, vjust = 0, size = 3, colour = "black", nudge_y = 0) +
  ggtitle("Connecticut New Covid-19 Cases (with 7-day rolling average)")

p_daily_deaths <- ggplot(data = dph_total %>% 
              # filter(county == "Fairfield") %>% 
              mutate(wday  = factor(weekdays(date),
                                  levels = c("Monday", "Tuesday",
                                             "Wednesday", "Thursday", "Friday",
                                             "Saturday","Sunday"))), 
            aes(x = date, y = new_deaths)) +
  geom_col(fill = "gray95", colour = "gray90") + 
  # scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.7, option = "viridis") + 
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", date_labels = "%b %d") +
  geom_line(aes(y = rnew_deaths), size = 1) +
  theme_minimal() + theme(legend.position = "none") +
  xlab(NULL) + ylab("Deaths Per Day") +
  geom_text(data = exec_orders %>% filter(date <= max(dph_total$date, na.rm = TRUE)), 
            aes(y = (rnew_deaths + 75) * fudge , label = label),
            hjust = 0.5, vjust = 0, size = 3, colour = "black", nudge_y = 0) +
  ggtitle("Connecticut Covid-19 Daily Deaths (with 7-day rolling average)")

r_data_ct <- rt_live %>% filter(region == "CT")
r_data_az <- rt_live %>% filter(region == "AZ")
# colors from colorbrewer:  https://colorbrewer2.org/#type=diverging&scheme=PRGn&n=3
p_rt_live <- ggplot(data = r_data_ct, aes(x = date, y = mean)) +
  geom_ribbon(aes(ymin = lower_90, ymax = upper_90), fill = "grey92") +
  geom_ribbon(data = r_data_az, aes(ymin = lower_90, ymax = upper_90), fill = "grey92") +
  geom_ribbon(aes(ymin = lower_50, ymax = upper_50), fill = "grey83") +
  geom_ribbon(data = r_data_az, aes(ymin = lower_50, ymax = upper_50), fill = "grey83") +
  geom_line(aes(colour = above_one)) + theme_minimal() + 
  geom_hline(yintercept = 1, colour = "gray") +
  geom_line(data = r_data_az, aes(colour = above_one)) + theme_minimal() + 
  geom_text(data = rt_live %>% filter(region %in% c("CT", "AZ"), date == min(date)),
            aes(label = state), hjust = 1, size = 3, nudge_x = -2) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle=0, vjust = 0.5)) +
  scale_color_manual(values = c(">= 1" = "#7fbf7b", "< 1" = "#af8dc3")) +
  scale_y_continuous(limits = c(0.5, 1.5), breaks = c(0.5, 1, 1.5)) +
                     #minor_breaks = seq(0.2, 1.4, 0.2)) +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", 
               date_labels = "%b %d", 
               limits = c(min(dph_total$date), max(dph_total$date))) +
  xlab(NULL) + ylab(expression("R"[t])) +
  # see: http://strata.uga.edu/8370/rtips/subscriptsSuperscripts.html
  labs(title = expression(paste("Trend in ", R[t], " (as computed by https://rt.live/)")))
```
```{r hospitalizations, echo = FALSE, out.width = "70%", fig.align = "center"}
p_hospitalizations <- ggplot(data = dph_total %>% filter(!is.na(hospital)),
                    aes(x = date, y = hospital)) +
  geom_line() + xlab(NULL) + # geom_point() + 
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", 
               date_labels = "%b %d", 
               limits = c(min(dph_total$date), NA)) +
  theme_minimal() + scale_y_continuous(labels = comma) +
  theme(legend.position = "none") +
  ylab("in hospital") +
  labs(
    title = "COVID19 Patients Currently in Hospital",
    subtitle = NULL,
    caption = "Source: Connecticut Hospital Association, via CT DPH"
  )
```
```{r geofacets, echo = FALSE}
p_rt_map <- ggplot(data = rt_live %>% select(-state) %>% 
         rename(state = region), aes(x = date, y = mean)) + 
  xlab(NULL) + ylab(NULL) + 
  theme_bw() + 
  geom_ribbon(aes(ymin = lower_90, ymax = upper_90), fill = "grey92") +
  geom_ribbon(aes(ymin = lower_50, ymax = upper_50), fill = "grey88") +
  geom_hline(yintercept = 1, colour = "grey50") +
  geom_line(aes(colour = above_one)) + theme(legend.position = "none") +
  scale_color_manual(values = c(">= 1" = "#7fbf7b", "< 1" = "#af8dc3")) +
  scale_y_continuous(limits = c(0.5, 1.5), breaks = c(0.5, 1, 1.5)) +
                     #minor_breaks = seq(0.2, 1.4, 0.2)) +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", 
               date_labels = "%b %d") + 
  theme(axis.title.y = element_text(angle=0, vjust = 0.5, size = 6)) +
  theme(axis.text.x = element_text(size=5)) +
  # labs(title = paste0("Trends in ", expression("R"[t]), "(as computed by https://rt.live/)")) +
  labs(title = expression(paste("Trend in ", R[t], " by State (as computed by https://rt.live/)"))) +
  facet_geo(~ state)

```
## Tracking the Virus in Connecticut

This post focuses on what is happening recently. The aim is to
use the recent trend to get some idea of the near future. The first
plot will show the average number of new cases reported each day along
with a line that displays the rolling seven-day average. Typically there 
are day-of-the-week effects in the reporting so it's best to focus
on the seven-day average.

The plot shows the history of new cases and also shows the 
actions required by some of Governor Lamont's executive orders. 
<br/><br/><br/>
```{r p-new-cases, echo = FALSE, fig.cap = "New Cases"}
print(p_new_cases)

```
<br/><br/><br/>
```{r p-rt-live, echo = FALSE, fig.asp = 0.3, fig.cap = "Estimated R<sub>t</sub>"}
print(p_rt_live)
```
<br/><br/><br/>
```{r p-daily-deaths, echo = FALSE, fig.cap = "Daily Deaths"}
print(p_daily_deaths)

```
<br/><br/><br/>
```{r p-hospitalizations, echo = FALSE, fig.cap = "Hospitalizations for Covid-19"}
print(p_hospitalizations)
```
<br/><br/>
The number of new cases peaked in mid April and then began to
decrease rapidly, although not as rapidly as it increased. 
A key issue is how fast epidemic is expanding or contracting.
One indicator of that is to estimate the
parameter R<sub>t</sub>, the average number of people 
who become infected by an infectious person. 
When R<sub>t</sub> is greater than 1 the number of daily new cases is
increasing, If R<sub>t</sub> is less than 1, it is decreasing.
See [this site](https://rt.live/) for
an estimate by state of the effective value for 
R<sub>t</sub>. The site
is described in [this article](https://www.vox.com/recode/2020/4/21/21227855/coronavirus-spreading-by-state-instagram-effective-reproduction-rate) at Vox.com.

Note that as of June (and a month
previous), Connecticut, New York, and New Jersey have among the lowest
values for R<sub>t</sub> in the country. This is good news for us
in Connecticut. I have included R<sub>t</sub> for Arizona as a contrast with
a state that has lately shown [signs of a growing outbreak](https://www.fox10phoenix.com/news/arizona-medical-experts-alarmed-over-surge-in-covid-19-cases).
The goal in Connecticut and in the entire New York City area is to
keep R<sub>t</sub> well under 1. As of `r max(rt_live$date)` R<sub>t</sub> is
`r round(rt_live$mean[(rt_live$date == max(rt_live$date)) & (rt_live$region == "CT")], 2)`. On the other hand, in Arizona R<sub>t</sub> is about `r round(rt_live$mean[(rt_live$date == max(rt_live$date)) & (rt_live$region == "AZ")], 2)`.

The next display shows the trend of R<sub>t</sub> in each state
arranged according to geographic position in the US. Remember
that R<sub>t</sub> is a measure of the rate of increase. When R<sub>t</sub> (indicated by the blue line) is greater than 1
(indicated by a dark gray line) the number of daily new cases is
increasing, If R<sub>t</sub> is less than 1, it is decreasing.
<br/><br/><br/>
```{r p-rt-map, echo = FALSE, fig.cap = "Estimate of R<sub>t</sub> by State"}
print(p_rt_map)
```
<br/><br/><br/>

## Deaths by Age

```{r deaths-by-age, echo = FALSE, out.width = '50%', fig.show = 'hold', fig.cap = "Relationship of Age to Covid-19 Deaths andCases"}
ggplot(data = dph_age %>% filter(date == max(date)), 
       aes(x = agegroups, y = deaths)) + 
  geom_col() + 
  labs(title = "Covid-19 Deaths in Connecticut by Age") +
  xlab("Age Groups") + ylab(NULL) +
  scale_y_continuous(labels = comma) +
  theme_minimal() + theme(panel.grid.major.x = element_blank())
# ```
# ```{r deaths-per-100k, echo = FALSE}
ggplot(data = dph_age %>% filter(date == max(date)), 
       aes(x = agegroups, y = as.numeric(per_100k))) + 
  geom_col() + 
  labs(title = "Covid-19 Cases by Age in Connecticut per 100K Population") +
  xlab("Age Groups") + ylab("Cases per 100K of Population") +
  scale_y_continuous(labels = comma) +
  theme_minimal() + theme(panel.grid.major.x = element_blank())
```

## Nursing Homes

```{r nursing-homes, echo = FALSE}
nursing_home_totals <- dph_nursing_cases %>% filter(date == max(date)) %>% 
  summarise(cases = sum(nh_cases, na.rm = TRUE), deaths = sum(nh_deaths, na.rm = TRUE))
```
Nursing homes have been hard hit by the epidemic. It's hard to control  coronavirus outbreak
in a group living situation, which includes prisons and cruises ship. On top of that,
Covid-19 is especially serious for the elderly. The combination in nursing homes
has been deadly. There have been `r comma(nursing_home_totals$deaths)` deaths in Connecticut 
Nursing Homes which makes `r paste0(round((nursing_home_totals$deaths/dph_total$deaths[dph_total$date == max(dph_total$date)]) * 100, 1), "%")` of the total of `r comma(dph_total$deaths[dph_total$date == max(dph_total$date)])` deaths from Covid-19.


## Variations by County in Connecticut

```{r county-trends, echo = FALSE}
county_centroid <- st_centroid(county_geometries) %>%
  left_join(dph_counties %>% filter(date == max(date, na.rm = TRUE)) %>%
              select(county, rnew_cases_per100k, cases, deaths, rnew_deaths_per100k), by = "county")
county_max <- dph_counties %>% group_by(county) %>% summarise(max = max(rnew_cases_per100k, na.rm = TRUE)) %>% arrange(desc(max))
for_plot1 <- dph_counties %>% 
  select(date, county, `new cases` = rnew_cases_per100k, deaths = rnew_deaths_per100k) %>% 
  mutate(county = factor(county, levels = county_max$county)) %>% 
  pivot_longer(cols = c(`new cases`, deaths)) %>% 
  mutate(the_measure = factor(name, levels = c("new cases", "deaths")))
p_county_per100k <- ggplot(data = for_plot1, aes(x = date, y = value * 100)) +
  geom_line(aes(colour = county)) +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", 
               date_labels = "%b %d") + 
  theme_minimal() + 
  xlab(NULL) + ylab("Per 100K Population") +
  facet_wrap(~ the_measure)
  # facet_wrap(~ name, scales = "free_y")
for_plot2 <- dph_counties %>% 
  select(date, county, cases = rcases, deaths = rdeaths) %>% 
  mutate(county = factor(county, levels = county_max$county)) %>% 
  pivot_longer(cols = c(cases, deaths)) %>% 
  mutate(the_measure = factor(name, levels = c("new cases", "deaths")))
p_county_cum <- ggplot(data = for_plot2, aes(x = date, y = value)) +
  geom_line(aes(colour = county)) +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", 
               date_labels = "%b %d") + 
  theme_minimal() + 
  xlab(NULL) + ylab("Cumulative Count (7-day rolling average)") +
  # facet_wrap(~ the_measure)
  facet_wrap(~ name, scales = "free_y")

p_county_map_new_cases <- ggplot() +
  geom_sf(data = county_geometries %>%
            left_join(dph_counties %>% filter(date == max(date, na.rm = TRUE)) %>%
                        select(county, rnew_cases_per100k),
                      by = "county") %>% 
            mutate(`new per 100k` = rnew_cases_per100k),
          # need aes(geometry = geometry) because joined sf to a tibble and so lost sf
          aes(fill = `new per 100k`, geometry = geometry)) +
  scale_fill_gradient(low = "white", high = "grey") +
  geom_sf_text(data = county_centroid, aes(label = paste0(county, "\n", round(rnew_cases_per100k, 1), " per 100K")), color = "black", size = 3) +
  coord_sf(datum = NA, label_axes = "----") +
  xlab("") + ylab("") + theme_minimal()  +
  # theme(legend.position = "none") +
  labs(title = "New Cases by Connecticut County",
       subtitle = "shading shows new cases per 100K of population as of most recent report",
       # fill = "cases per 1K",
       caption = "Sources: CT DPH, US Census, tidycensus package")

p_county_map_cum_cases <- ggplot() +
  geom_sf(data = county_geometries %>%
            left_join(dph_counties %>% filter(date == max(date, na.rm = TRUE)) %>%
                        select(county, cases, deaths),
                      by = "county") %>% 
            mutate(`cases per 100k` = cases / (total_pop / 100000)),
          # need aes(geometry = geometry) because joined sf to a tibble and so lost sf
          aes(fill = `cases per 100k`, geometry = geometry)) +
  scale_fill_gradient(low = "white", high = "grey") +
  geom_sf_text(data = county_centroid, aes(label = paste0(county, "\n", comma(cases, accuracy = 1.0), " cases\n", comma(deaths, accuracy = 1.0), " deaths")), color = "black", size = 3) +
  coord_sf(datum = NA, label_axes = "----") +
  xlab("") + ylab("") + theme_minimal()  +
  # theme(legend.position = "none") +
  labs(title = "Cumulative Cases and Deaths by Connecticut County",
       subtitle = "shading shows total cases per 100K of population",
       # fill = "cases per 1K",
       caption = "Source: CT DPH, US Census, tidycensus package")

```
Next I'll look at variations within Connecticut. Below are two maps
showing the eight counties in Connecticut. The first map in Figure \@ref(fig:county-maps) shows new cases 
per 1K of population. (Actually it show the rolling average the last seven days
in order to smooth out varitions in day to day reporting.) The second map shows
the cumulative total of cases. So the first map is an indication of the recent prevalence of the
virus and the second map relates more to the total impact of the disease over the
full course of the epidemic to date.

Connecticut is part of the epidemic's surge in the New York area and the magnitude
of the epidemic has been greater in the counties closer to New York City. Back in March cases
first appeared in Fairfield County and then spread to Litchfield, New Haven, and Hartford.
More recently Hartford has been the area that we have to watch and even the most eastern
counties in the state.
<br/><br/><br/>
```{r county-maps, echo = FALSE, fig.cap = "County Maps of New and of Cumulative Cases", out.width = '50%', fig.show = 'hold'}
print(p_county_map_cum_cases)
print(p_county_map_new_cases)
```
```{r town-categories, echo = FALSE}
by_category <- dph_towns %>% 
  combine_dph_towns(category)
p_by_category <- ggplot(data = by_category,
                        aes(x = date, y = rnew_cases_per100k)) +
  geom_line() +
  facet_wrap(~ category)

```

### By County and Type of Town

There are 169 towns in Connecticut, which makes it hard to eamine variations by town. To help
examine variations within counties, I'll use a typology of towns called 
[The Five Connecticuts](https://ctdatahaven.org/sites/ctdatahaven/files/UConnCPR%20Changing%20Demographics-5%20CTs%202004.pdf)
that divides towns into five categories based on census variables: Urban Core, Urban Periphery,
Suburban, Rural, and a fifth category for Wealthy Suburbs used for some towns in Fairfield County. 
Adjusting for population, the number of cases has been greater in the counties closer
to New York City. 
In the figure below I'll fold the wealthy towns into the Suburban category. I have excluded
Montville and Somers because the prisons in those towns complicates interpretation of the
town statistics.

Counties are laid out from top to bottom and the counties closer to New York are toward the top. 
The columns show the four categories. Urban towns have been hit harder than suburban and rural
towns. While density may have some effect on case rates, the report [Towards Health Equity in Connecticut: The Role of Social Inequality and the Impact of COVID-19](https://ctdatahaven.org/reports/towards-health-equity-connecticut)
by DataHaven documents the unequal impact of the epidemic in Connecticut related to social inequality.

```{r county-by-category, echo = FALSE, fig.asp = 1.2}

# create 
for_chart <- dph_towns %>% filter(!(town %in% c("Montville", "Somers"))) %>% 
  mutate(category = fct_recode(category, Suburban = "Wealthy")) %>% 
  combine_dph_towns(county, category) %>% 
  mutate(rnew_cases_per100k = rnew_cases_per100k,
    county = factor(county,
      levels = c("Litchfield", "Fairfield", "Hartford", "New Haven", "Middlesex", "Tolland", "New London", "Windham")),
    for_label = paste0(towns, ifelse(towns > 1, " towns", "town"), ", pop=", comma(total_pop, accuracy = 1)))
ggplot(data = for_chart,
                        aes(x = date, y = rnew_cases_per100k)) +
  geom_line() +
  # facet_grid(county ~ category, nrow = 8, drop = FALSE) +
  facet_grid(county ~ category, drop = FALSE) +
  geom_text(data = for_chart %>% filter(date == min(date)),
            x = -Inf, y = Inf,
            aes(label = for_label), hjust = 0, vjust = 1, size = 3) +
  ylab("New Cases per 100K") + xlab(NULL) +
  labs(title = "New Cases by County and by Type of Town Per Population",
       subtitle = "(7-day rolling average of new cases)",
       caption = '"Wealthy" towns are included with "Suburban".\nSomers and Montville are omitted because prison data may dominate town data.',
       tag = NULL)

```


## Methodology and Sources

#### Counts from Connecticut Department of Public Health
The charts shows the counts of cases and deaths as they are reported by
the Connecticut Department of Public Health. Reports sometimes end up in
batches. There is a day of the week effect, with fewer cases on weekends.
Some of the unusually large peaks and valleys in these charts are due
to reporting process. The DPH has also begun producing reports based
on the date a sample was taken for a test and for the date of
death. That's a more accurate way to look at the change over time, but it
means the data for the most recent days are hard to interpret because
some data may be "in process" and not yet reported. In these charts I 
have used "date of report" rather than "date of sample taken" or "date of death"
because that gives me all of the recent data that is available
and because that is what most other data projects (such as the Covid-19 Project
or the New York Times) have been using.



