---
title: Tracking Covid-19 in Connecticut
author: John Goldin
date: '2020-06-12'
slug: []
categories:
  - COVID19
tags:
  - COVID19
type: post
subtitle: ''
image: ''
draft: yes
output:
#  blogdown::html_page:
  bookdown::html_document2:
    toc: yes
    fig_width: 7
    out.width: "70%" 
    fig.align: "center"
    fig.asp: 0.618
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, out.width = "100%",
                      knitr.table.format = "html")
# knitr.table.format is for kable. Needed this to get blogdown table to format correctly.
```
```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse, quietly = TRUE)
library(knitr, quietly = TRUE)
library(scales, quietly = TRUE)
library(kableExtra, quietly = TRUE)
library(sparkline)
library(geofacet) # https://hafen.github.io/geofacet/
library(sf, quietly = TRUE)
options(tigris_use_cache = TRUE)
library(tigris, quietly = TRUE)
library(tidycensus, quietly = TRUE)
library(viridis, quietly = TRUE)
library(ggrepel)
options(tigris_class = "sf")
library(lubridate, quietly = TRUE)
library(janitor)
library(fuzzyjoin) # for interval_left_join
library(broom)
library(RSocrata)
library(RcppRoll)
```
```{r `load_basic_data`, echo = FALSE, message = FALSE}
path_to_post <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/content/post/2020-03-29-covid19-cases-in-connecticut/"
path_to_static_for_this_post <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/static/post/2020-06-12-tracking-covid-19-in-connecticut"
path_to_ctcorona <- "~/Documents/R_local_repos/ctcorona/data/"
load(paste0(path_to_ctcorona, "dph_datasets.RData"))
load(paste0(path_to_ctcorona, "census_population.RData"))

# from https://www.ctdatahaven.org/sites/ctdatahaven/files/UConnCPR%20Changing%20Demographics-5%20CTs%202004.pdf
# The Changing Demographics of Connecticut — 1990 to 2000. by Center for Population Research  May. 31, 2004
Five_Connecticuts <- read_delim(paste0(path_to_ctcorona, "Five_Connecticuts.txt"),"\t", escape_double = FALSE, trim_ws = TRUE) %>%
  mutate(category = factor(category, levels = c("Urban Core", "Urban Periphery",
                           "Wealthy", "Suburban", "Rural")))
load(paste0(path_to_ctcorona, "dph_datasets.RData"))
# load effective R from https://rt.live/
rt_live <- read_csv("https://d14wlfuexuxgcm.cloudfront.net/covid/rt.csv") %>% 
  mutate(state = case_when(
    region == "CT" ~ "Connecticut",
    region == "AZ" ~ "Arizona"
  ))

```
The [post I created back in March](https://johngoldin.com/post/2020-03-29-covid19-cases-in-connecticut/) 
aimed to track the growth of Covid-19 as it hit Connecticut.
Back then (at the end of March) I did not realize that 
the greater New York area (including Connecticut)
would be the center of the first wave in the US. At that time what I was reading
focused on the nature of exponential growth of an epidemic. We were urged to
"flatten the curve" and there was an emphasis on the shape of the growth of
cumulative cases and deaths caused by the virus. I have been upading the figures
in that post each evening. During that time the [Connecticut data portal](https://data.ct.gov/stories/s/COVID-19-data/wa3g-tfvc/)
has expanded considerably, and I've built [some R code](https://github.com/johngoldin/ctcorona/blob/master/R/daily_ct_stats.R) 
that let's me easily update
that post. I'll use the same data infrastructure for this post as well.

```{r 'new-cases-plot', echo = FALSE, eval = TRUE}
## This imaged will be saved in the /post/*_files/ directory
## Use echo = FALSE if you want to hide the code for making the plot
exec_orders <- tibble(
  date = as.Date(c("2020-03-10", "2020-03-17", "2020-03-23", "2020-03-26", "2020-04-11", "2020-04-20", "2020-05-20", "2020-06-01", "2020-06-17")),
  label = c("prohibit large\ngatherings", "cancel\nclasses", "restrict\nbusiness", "5 or\nfewer",
            "covid-19\ndeath\nchanged", "masks\nrequired",
            "phase 1\nreopen", "hair\nreopen", "phase 2\nreopen"),
  fudge = c(1, 0.4, 0.7, 0.9, 0.8, 0.9, 0.8, 0.8, 1)
) %>% 
  left_join(dph_total %>% select(date, rnew_cases, rnew_deaths, cases, deaths))

# try a version with understated bars
p_new_cases <- ggplot(data = dph_total %>% 
              # filter(county == "Fairfield") %>% 
              mutate(wday  = factor(weekdays(date),
                                  levels = c("Monday", "Tuesday",
                                             "Wednesday", "Thursday", "Friday",
                                             "Saturday","Sunday"))), 
            aes(x = date, y = new_cases)) +
  geom_col(fill = "gray95", colour = "gray90") + 
  # scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.7, option = "viridis") + 
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", date_labels = "%b %d") +
  geom_line(aes(y = rnew_cases), size = 1) +
  theme_minimal() + theme(legend.position = "none") +
  xlab(NULL) + ylab("New Cases Per Day") +
  geom_text(data = exec_orders %>% filter(date <= max(dph_total$date, na.rm = TRUE)), 
            aes(y = (rnew_cases + 500) * fudge , label = label),
            hjust = 0.5, vjust = 0, size = 3, colour = "black", nudge_y = 0) +
  ggtitle("Connecticut New Covid-19 Cases (with 7-day rolling average)")

p_daily_deaths <- ggplot(data = dph_total %>% 
              # filter(county == "Fairfield") %>% 
              mutate(wday  = factor(weekdays(date),
                                  levels = c("Monday", "Tuesday",
                                             "Wednesday", "Thursday", "Friday",
                                             "Saturday","Sunday"))), 
            aes(x = date, y = new_deaths)) +
  geom_col(fill = "gray95", colour = "gray90") + 
  # scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.7, option = "viridis") + 
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", date_labels = "%b %d") +
  geom_line(aes(y = rnew_deaths), size = 1) +
  theme_minimal() + theme(legend.position = "none") +
  xlab(NULL) + ylab("Deaths Per Day") +
  geom_text(data = exec_orders %>% filter(date <= max(dph_total$date, na.rm = TRUE)), 
            aes(y = (rnew_deaths + 75) * fudge , label = label),
            hjust = 0.5, vjust = 0, size = 3, colour = "black", nudge_y = 0) +
  ggtitle("Connecticut Covid-19 Daily Deaths (with 7-day rolling average)")

r_data_ct <- rt_live %>% filter(region == "CT")
r_data_az <- rt_live %>% filter(region == "AZ")
p_rt_live <- ggplot(data = r_data_ct, aes(x = date, y = mean)) +
  geom_ribbon(aes(ymin = lower_90, ymax = upper_90), fill = "grey92") +
  geom_ribbon(data = r_data_az, aes(ymin = lower_90, ymax = upper_90), fill = "grey92") +
  geom_ribbon(aes(ymin = lower_50, ymax = upper_50), fill = "grey83") +
  geom_ribbon(data = r_data_az, aes(ymin = lower_50, ymax = upper_50), fill = "grey83") +
  geom_line() + theme_minimal() + 
  geom_hline(yintercept = 1, colour = "gray") +
  geom_line(data = r_data_az) + theme_minimal() + 
  geom_text(data = rt_live %>% filter(region %in% c("CT", "AZ"), date == min(date)),
            aes(label = state), hjust = 1, size = 3) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle=0, vjust = 0.5)) +
  scale_y_continuous(limits = c(0.5, 1.5), breaks = c(0.5, 1, 1.5)) +
                     #minor_breaks = seq(0.2, 1.4, 0.2)) +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", 
               date_labels = "%b %d", 
               limits = c(min(dph_total$date), max(dph_total$date))) +
  xlab(NULL) + ylab(expression("R"[t])) +
  # see: http://strata.uga.edu/8370/rtips/subscriptsSuperscripts.html
  labs(title = expression(paste("Trend in ", R[t], " (as computed by https://rt.live/)")))
```
```{r hospitalizations, echo = FALSE, out.width = "70%", fig.align = "center"}
p_hospitalizations <- ggplot(data = dph_total %>% filter(!is.na(hospital)),
                    aes(x = date, y = hospital)) +
  geom_line() + xlab(NULL) + # geom_point() + 
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", 
               date_labels = "%b %d", 
               limits = c(min(dph_total$date), NA)) +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("in hospital") +
  labs(
    title = "COVID19 Patients Currently in Hospital",
    subtitle = NULL,
    caption = "Source: Connecticut Hospital Association, via CT DPH"
  )
```
```{r geofacets, echo = FALSE}
p_rt_map <- ggplot(data = rt_live %>% select(-state) %>% 
         rename(state = region), aes(x = date, y = mean)) + 
  xlab(NULL) + ylab(NULL) + 
  theme_bw() + 
  geom_ribbon(aes(ymin = lower_90, ymax = upper_90), fill = "grey92") +
  geom_ribbon(aes(ymin = lower_50, ymax = upper_50), fill = "grey88") +
  geom_hline(yintercept = 1, colour = "grey50") +
  geom_line(colour = "darkblue") +
   scale_y_continuous(limits = c(0.5, 1.5), breaks = c(0.5, 1, 1.5)) +
                     #minor_breaks = seq(0.2, 1.4, 0.2)) +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", 
               date_labels = "%b %d") + 
  theme(axis.title.y = element_text(angle=0, vjust = 0.5, size = 6)) +
  theme(axis.text.x = element_text(size=5)) +
  # labs(title = paste0("Trends in ", expression("R"[t]), "(as computed by https://rt.live/)")) +
  labs(title = expression(paste("Trend in ", R[t], " by State (as computed by https://rt.live/)"))) +
  facet_geo(~ state)

```
### Tracking the Virus in Connecticut

This post focuses on what is happening recently. The aim is to
use the recent trend to get some idea of the near future. The first
plot will show the average number of new cases reported each day along
with a line that displays the rolling seven-day average. Typically there 
are day-of-the-week effects in the reporting so it's best to focus
on the seven-day average.

The plot show the history of new cases and also shows the 
actions in some of Governor Lamont's executive orders. 

```{r p-new-cases, echo = FALSE, fig.cap = "New Cases"}
print(p_new_cases)

```
```{r p-rt-live, echo = FALSE, fig.asp = 0.3, fig.cap = "Estimated R<sub>t</sub>"}
print(p_rt_live)
```
```{r p-daily-deaths, echo = FALSE, fig.cap = "Daily Deaths"}
print(p_daily_deaths)

```
```{r p-hospitalizations, echo = FALSE, fig.cap = "Hospitalizations for Covid-19"}
print(p_hospitalizations)
```

The number of new cases peaked in mid April and then began to
decrease rapidly, although not as rapidly as it increased. 
See [this site](https://rt.live/) for
an estimate by state of the effective value for 
R<sub>t</sub>, the average number of people 
who become infected by an infectious person. The site
is described in [this article](https://www.vox.com/recode/2020/4/21/21227855/coronavirus-spreading-by-state-instagram-effective-reproduction-rate) at Vox.com.

Note that as of June (and a month
previous), Connecticut, New York, and New Jersey have among the lowest
values for R<sub>t</sub> in the country. This is good news for us
in Connecticut. I have included R<sub>t</sub> for Arizona as a contrast with
a state that has lately shown [signs of a growing outbreak](https://www.fox10phoenix.com/news/arizona-medical-experts-alarmed-over-surge-in-covid-19-cases).
The goal in Connecticut and in the entire New York City area is to
keep R<sub>t</sub> well under 1. As of `r max(rt_live$date)` R<sub>t</sub> is
`r round(rt_live$mean[(rt_live$date == max(rt_live$date)) & (rt_live$region == "CT")], 2)`. On the other hand, in Arizona R<sub>t</sub> is about `r round(rt_live$mean[(rt_live$date == max(rt_live$date)) & (rt_live$region == "AZ")], 2)`.

The next display shows the trend of R<sub>t</sub> in each state
arranged according to geographic position in the US. Remember
that R<sub>t</sub> is a measure of the rate of increase. When R<sub>t</sub> (indicated by the blue line) is greater than 1
(indicated by a dark gray line) the number of daily new cases is
increasing, If R<sub>t</sub> is less than 1, it is decreasing.

```{r p-rt-map, echo = FALSE, fig.cap = "Estimate of R<sub>t</sub> by State"}
print(p_rt_map)
```

### Variations by County in Connecticut

```{r county-trends, echo = FALSE}
county_centroid <- st_centroid(county_geometries) %>%
  left_join(dph_counties %>% filter(date == max(date, na.rm = TRUE)) %>%
              select(county, rnew_cases_per1k, cases, deaths, rnew_deaths_per1k), by = "county")
county_max <- dph_counties %>% group_by(county) %>% summarise(max = max(rnew_cases_per1k, na.rm = TRUE)) %>% arrange(desc(max))
for_plot1 <- dph_counties %>% 
  select(date, county, `new cases` = rnew_cases_per1k, deaths = rnew_deaths_per1k) %>% 
  mutate(county = factor(county, levels = county_max$county)) %>% 
  pivot_longer(cols = c(`new cases`, deaths)) %>% 
  mutate(the_measure = factor(name, levels = c("new cases", "deaths")))
p_county_per100k <- ggplot(data = for_plot1, aes(x = date, y = value * 100)) +
  geom_line(aes(colour = county)) +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", 
               date_labels = "%b %d") + 
  theme_minimal() + 
  xlab(NULL) + ylab("Per 100K Population") +
  facet_wrap(~ the_measure)
  # facet_wrap(~ name, scales = "free_y")
for_plot2 <- dph_counties %>% 
  select(date, county, cases = rcases, deaths = rdeaths) %>% 
  mutate(county = factor(county, levels = county_max$county)) %>% 
  pivot_longer(cols = c(cases, deaths)) %>% 
  mutate(the_measure = factor(name, levels = c("new cases", "deaths")))
p_county_cum <- ggplot(data = for_plot2, aes(x = date, y = value)) +
  geom_line(aes(colour = county)) +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", 
               date_labels = "%b %d") + 
  theme_minimal() + 
  xlab(NULL) + ylab("Cumulative Count (7-day rolling average)") +
  # facet_wrap(~ the_measure)
  facet_wrap(~ name, scales = "free_y")

p_county_map_new_cases <- ggplot() +
  geom_sf(data = county_geometries %>%
            left_join(dph_counties %>% filter(date == max(date, na.rm = TRUE)) %>%
                        select(county, rnew_cases_per1k),
                      by = "county"),
          # need aes(geometry = geometry) because joined sf to a tibble and so lost sf
          aes(fill = rnew_cases_per1k, geometry = geometry)) +
  scale_fill_gradient(low = "white", high = "grey") +
  geom_sf_text(data = county_centroid, aes(label = paste0(county, "\n", round(rnew_cases_per1k, 3), " per 1K")), color = "black", size = 3) +
  coord_sf(datum = NA, label_axes = "----") +
  xlab("") + ylab("") + theme_minimal()  +
  # theme(legend.position = "none") +
  labs(title = "New Cases by Connecticut County",
       subtitle = "per 1K of population as of recent report date",
       fill = "cases per 1K",
       caption = "Source: US Census, tidycensus package")

p_county_map_cum_cases <- ggplot() +
  geom_sf(data = county_geometries %>%
            left_join(dph_counties %>% filter(date == max(date, na.rm = TRUE)) %>%
                        select(county, cases),
                      by = "county"),
          # need aes(geometry = geometry) because joined sf to a tibble and so lost sf
          aes(fill = cases / (total_pop / 1000), geometry = geometry)) +
  scale_fill_gradient(low = "white", high = "grey") +
  geom_sf_text(data = county_centroid, aes(label = paste0(county, "\n", comma(cases), " cases")), color = "black", size = 3) +
  coord_sf(datum = NA, label_axes = "----") +
  xlab("") + ylab("") + theme_minimal()  +
  # theme(legend.position = "none") +
  labs(title = "Cumulative Cases by Connecticut County",
       subtitle = "cases per 1K of population for most recent report date",
       fill = "cases per 1K",
       caption = "Source: US Census, tidycensus package")
```
Next I'll look at variations within Connecticut. Below are two maps
showing the eight counties in Connecticut. The first map in Figure \@ref(fig:county-maps) shows new cases 
per 1K of population. (Actually it show the rolling average the last seven days
in order to smooth out varitions in day to day reporting.) The second map shows
the cumulative total of cases. So the first map is an indication of the recent prevalence of the
virus and the second map relates more to the total impact of the disease over the
full course of the epidemic to date.

Connecticut is part of the epidemic's surge in the New York area and the magnitude
of the epidemic has been greater in the counties closer to New York City. Back in March cases
first appeared in Fairfield County and then spread to Litchfield, New Haven, and Hartford.
More recently Hartford has been the area that we have to watch and even the most eastern
counties in the state.

```{r county-maps, echo = FALSE, fig.cap = "County Maps of New and of Cumulative Cases", out.width = '50%', fig.show = 'hold'}
print(p_county_map_new_cases)
#```
#```{r county_maps_cum_cases, echo = FALSE, fig.cap = "County Map Cumulative Cases"}
print(p_county_map_cum_cases)
```



### Methodology and Sources

#### Counts from Connecticut Department of Public Health
The charts shows the counts of cases and deaths as they are reported by
the Connecticut Department of Public Health. Reports sometimes end up in
batches. There is a day of the week effect, with fewer cases on weekends.
Some of the unusually large peaks and valleys in these charts are due
to reporting process. The DPH has also begun producing reports based
on the date a sample was taken for a test and for the date of
death. That's a more accurate way to look at the change over time, but it
means the data for the most recent days are hard to interpret because
some data may be "in process" and not yet reported. In these charts I 
have used "date of report" rather than "date of sample taken" or "date of death"
because that gives me all of the recent data that is available
and because that is what most other data projects (such as the Covid-19 Project
or the New York Times) have been using.



