---
title: Tracking Covid-19 in Connecticut
author: John Goldin
date: '2020-06-12'
slug: []
categories:
  - COVID19
tags:
  - COVID19
type: post
subtitle: ''
image: ''
draft: yes
output:
  blogdown::html_page:
    toc: yes
    fig_width: 7
    fig_height: 5
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, out.width = "100%",
                      knitr.table.format = "html")
# knitr.table.format is for kable. Needed this to get blogdown table to format correctly.
```
```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse, quietly = TRUE)
library(knitr, quietly = TRUE)
library(scales, quietly = TRUE)
library(kableExtra, quietly = TRUE)
library(sparkline)
```
```{r `load_basic_data`, echo = FALSE, message = FALSE}
path_to_post <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/content/post/2020-03-29-covid19-cases-in-connecticut/"
path_to_ctcorona <- "~/Documents/R_local_repos/ctcorona/data/"
load(paste0(path_to_ctcorona, "dph_datasets.RData"))
load(paste0(path_to_ctcorona, "census_population.RData"))

# from https://www.ctdatahaven.org/sites/ctdatahaven/files/UConnCPR%20Changing%20Demographics-5%20CTs%202004.pdf
# The Changing Demographics of Connecticut — 1990 to 2000. by Center for Population Research  May. 31, 2004
Five_Connecticuts <- read_delim(paste0(path_to_ctcorona, "Five_Connecticuts.txt"),"\t", escape_double = FALSE, trim_ws = TRUE) %>%
  mutate(category = factor(category, levels = c("Urban Core", "Urban Periphery",
                           "Wealthy", "Suburban", "Rural")))
load(paste0(path_to_ctcorona, "dph_datasets.RData"))

```
The [post I created back in March](https://johngoldin.com/post/2020-03-29-covid19-cases-in-connecticut/) 
aimed to track the growth of Covid-19 as it hit Connecticut.
Back then (at the end of March) I didn't realize that 
the greater New York area (including Connecticut)
would be the center of the first wave in the US. At that time what I was reading
focused on the nature of exponential growth of an epidemic. We were urged to
"flatten the curve" and there was an emphasis on the shape of the growth of
cumulative cases and deaths caused by the virus. I have been upading the figures
in that post each evening. During tht time the [Connecticut data portal](https://johngoldin.com/post/2020-03-29-covid19-cases-in-connecticut/)
has expanded, and I've built [some R code](https://github.com/johngoldin/ctcorona/blob/master/R/daily_ct_stats.R) 
that let's me easily update
that post. I'll use the same data infrastructure for this post as well.

### Tracking the Virus

The focus on this post is what's happening recently. The first
plot will show the average number of new cases reported each day along
with a line that displays the rolling seven-day average. Typically there 
are day-of-the-week effects in the reporting so it's best to focus
on the seven-day average.

The plot show the history of new cases and also shows the 
actions in Governor Lamont's executive orders. 

The number of new cases peaked in mid April and then began to
decrease almost as rapidly.

```{r 'new_cases_plot', echo = FALSE, eval = TRUE}
## This imaged will be saved in the /post/*_files/ directory
## Use echo = FALSE if you want to hide the code for making the plot
exec_orders <- tibble(
  date = as.Date(c("2020-03-10", "2020-03-17", "2020-03-23", "2020-03-26", "2020-04-11", "2020-04-20", "2020-05-20", "2020-06-01", "2020-06-17")),
  label = c("prohibit large\ngatherings", "cancel\nclasses", "restrict\nbusiness", "5 or\nfewer",
            "covid-19\ndeath\nchanged", "masks\nequired",
            "phase 1\nreopen", "hair\nreopen", "phase 2\nreopen"),
  fudge = c(1, 0.4, 0.7, 0.9, 0.8, 0.9, 0.8, 0.8, 1)
) %>% 
  left_join(dph_total %>% select(date, rnew_cases, rnew_deaths, cases, deaths))

# try a version with understated bars
p_new_cases <- ggplot(data = dph_total %>% 
              # filter(county == "Fairfield") %>% 
              mutate(wday  = factor(weekdays(date),
                                  levels = c("Monday", "Tuesday",
                                             "Wednesday", "Thursday", "Friday",
                                             "Saturday","Sunday"))), 
            aes(x = date, y = new_cases)) +
  geom_col(fill = NA, colour = "gray") + 
  # scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.7, option = "viridis") + 
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", date_labels = "%b %d") +
  geom_line(aes(y = rnew_cases), size = 1) +
  theme_minimal() + theme(legend.position = "none") +
  xlab(NULL) + ylab("New Cases Per Day") +
  geom_text(data = exec_orders %>% filter(date <= max(dph_total$date, na.rm = TRUE)), 
            aes(y = (rnew_cases + 500) * fudge , label = label),
            hjust = 0.5, vjust = 0, size = 3, colour = "black", nudge_y = 0) +
  ggtitle("Connecticut New Covid-19 Cases (with 7-day rolling average)")

p_daily_deaths <- ggplot(data = dph_total %>% 
              # filter(county == "Fairfield") %>% 
              mutate(wday  = factor(weekdays(date),
                                  levels = c("Monday", "Tuesday",
                                             "Wednesday", "Thursday", "Friday",
                                             "Saturday","Sunday"))), 
            aes(x = date, y = new_deaths)) +
  geom_col(fill = NA, colour = "gray") + 
  # scale_fill_viridis(discrete = TRUE, begin = 0.5, end = 0.7, option = "viridis") + 
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", date_labels = "%b %d") +
  geom_line(aes(y = rnew_deaths), size = 1) +
  theme_minimal() + theme(legend.position = "none") +
  xlab(NULL) + ylab("Deaths Per Day") +
  geom_text(data = exec_orders %>% filter(date <= max(dph_total$date, na.rm = TRUE)), 
            aes(y = (rnew_deaths + 75) * fudge , label = label),
            hjust = 0.5, vjust = 0, size = 3, colour = "black", nudge_y = 0) +
  ggtitle("Connecticut Covid-19 Daily Deaths (with 7-day rolling average)")
print(p_new_cases)
print(p_daily_deaths)
```

```{r hospitalizations, echo = FALSE}
p_hospitalizations <- ggplot(data = dph_total %>% filter(!is.na(hospital)),
                    aes(x = date, y = hospital)) +
  geom_line() + xlab(NULL) + # geom_point() + 
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", 
               date_labels = "%b %d", limits = c(min(dph_total$date), NA)) +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("in hospital") +
  labs(
    title = "COVID19 Patients Currently in Hospital",
    subtitle = NULL,
    caption = "Source: Connecticut Hospital Association, via CT DPH"
  )
print(p_hospitalizations)
```


