---
author: 
date: "`r Sys.Date()`"
output: html_document
params: 
  county: "New Haven"
  # output_dir: "../2020-06-07-test-post-two"
  output_dir: "~/Dropbox/Programming/R_Stuff/can_i_blog_too/static/2020-06-22-tracking-covid-19-in-connecticut"
  return_url: "index.html"
title: "Covid-19 Summary for `r params$county` County"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# see this StackOverflow tip for info on how to set title via parameter:
#    https://stackoverflow.com/questions/31861569/setting-document-title-in-rmarkdown-from-parameters
```
```{r library, echo = FALSE, message = FALSE}
library(tidyverse)
library(sparkline)
library(kableExtra)
library(formattable)
library(sf)
library(viridis)
```
```{r setup_town_info, echo = FALSE}
path_to_post <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/content/post/2020-03-29-covid19-cases-in-connecticut/"
path_to_ctcorona <- "~/Documents/R_local_repos/ctcorona/data/"
path_to_post_june <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/content/post/2020-06-22-tracking-covid-19-in-connecticut/"
path_to_static_june <- "~/Dropbox/Programming/R_Stuff/can_i_blog_too/static/post/2020-06-22-tracking-covid-19-in-connecticut/"

load(paste0(path_to_ctcorona, "dph_datasets.RData"))
load(paste0(path_to_ctcorona, "census_population.RData"))
```

This table summarizes trends in each town in `r params$county` County. 
It is sorted by the category of town as
developed in the report [The Five Connecticuts](https://ctdatahaven.org/sites/ctdatahaven/files/UConnCPR%20Changing%20Demographics-5%20CTs%202004.pdf):
Urban Core, Urban Periphery, Wealthy, Suburban, and Rural.

For each town it shows the category, the population, 
the total number of cases to date, the number of recent cases per
100K of population (new cases in the last two weeks), total deaths, and the percentage of total deaths
in that town
attributed to nursing home residents.

Next there are two [sparklines](https://en.wikipedia.org/wiki/Sparkline)
that show the trend in daily new cases and daily deaths (based on 
14 day moving average). Cases and deaths are shown as a ratio to population
in the town. 

The purpose of this table is to make it easier to quickly scan
for trends in local towns.

```{r sparkline_ranges, echo  = FALSE}
county_cases_max <- max(dph_counties$rnew_cases_per100k, na.rm = TRUE) 
county_deaths_max <- max(dph_counties$rnew_deaths_per100k, na.rm = TRUE) 
town_cases_max <- max(dph_towns$rnew_cases_per100k, na.rm = TRUE) 
town_cases_max <- 1.0 * 100 # Somers is 2 and Montville is 1.06 
town_deaths_max <- max(dph_towns$rnew_deaths_per100k, na.rm = TRUE) 


# For the county sparklines, maximum range for cases is `r round(county_cases_max,1)`
# maximum range for deaths is `r round(county_deaths_max, 1)`
```

The town sparklines at the right show the trends over time.
The vertical scale is different for new cases and for deaths.
For new cases, the maximum height shown on the scale is `r round(town_cases_max, 1)` while 
for deaths, the maximum value shown is `r round(town_deaths_max, 1)`. For both new cases and
deaths, the sparkline shows the rolling seven day average of new cases or deaths. 



```{r, echo = FALSE, results = "asis"}
# THE ORIGINAL VERSION
# based on https://stackoverflow.com/questions/54244718/combining-sparkline-in-kableextra-table
# jQuery sparklines https://omnipotent.net/jquery.sparkline/#s-docs
county_sparklines <- dph_counties %>% 
  group_by(county) %>% 
  summarise(`daily cases` = spk_chr(rnew_cases_per100k, chartRangeMin = 0,
                                     chartRangeMax = county_cases_max),
            `daily deaths` = spk_chr(rnew_deaths_per100k, chartRangeMin = 0,
                                     chartRangeMax = county_deaths_max), .groups = "drop")

# county_summary <- dph_counties %>% 
#   filter(county == params$county)
# county_summary <- county_summary %>% 
#   # left_join(town_info %>% select(county, category, town)) %>% 
#   group_by(county) %>% 
#   # summarise(`max cases` = round(max(rnew_cases, na.rm = TRUE)),
#   summarise(`max cases` = (max(rnew_cases_per100k, na.rm = TRUE) %>% round(4)),
#     # `county cases` = spk_chr(rnew_cases_per100k, chartRangeMin = 0, 
#     #                        chartRangeMax = county_cases_max),
#     # `county deaths` = spk_chr(rnew_deaths_per100k, chartRangeMin = 0, 
#     #                        chartRangeMax = county_deaths_max), 
#     .groups = "drop") %>% 
#   # left_join(county_sparklines, by = "county") %>% 
#   # left_join(county_info %>% select(county, total_pop), by = "county") %>% 
#   # left_join(town_with_nursing %>% 
#   #             filter(date == max(date)) %>% 
#   #             select(town, nh_death_pct), by = "town") %>% 
#   mutate(nh_death_pct = NA_real_, category = "") %>% 
#   select(county, category, county, total_pop, nh_death_pct, `max cases`, `daily cases`,
#          `daily deaths`) %>% 
#   mutate(total_pop = comma(total_pop, digits = 0), 
#          nh_death_pct = percent(nh_death_pct, digits = 0)) %>% 
#   formattable::format_table(
#     # x = .,
#     # formatters = list(align=c("l")),
#     format = "html",
#     align = c("l", "l", "l", "l", "c", "c", "c", "c")
#   ) %>%
#   # kable_styling("striped", full_width = F) %>%
#   # kable_styling(full_width = F) %>%
#   # group_rows("group1", 1, 2) %>%
#   # group_rows("group1", 1, 2) %>%
#   # group_rows("group2", 3,3) %>%
#   htmltools::HTML() %>%
#   shiny::div() %>%
#   sparkline::spk_add_deps()


town_summary <- dph_towns %>%
  filter(county == params$county, #total_pop > 10000,
         !is.na(rnew_cases))  %>%
  # left_join(town_info %>% select(county, category, town)) %>% 
  group_by(county, category, town) %>% 
  # summarise(`max cases` = round(max(rnew_cases, na.rm = TRUE)),
  summarise(`total cases` = max(cases, na.rm = TRUE),
            `total deaths` = max(deaths, na.rm = TRUE),
            total_pop = scales::comma(max(total_pop, na.rm = TRUE)),
            
    `town cases` = spk_chr(rnew_cases_per100k, chartRangeMin = 0, 
                           chartRangeMax = town_cases_max),
    `town deaths` = spk_chr(rnew_deaths_per100k, chartRangeMin = 0, 
                           chartRangeMax = town_deaths_max), .groups = "drop") %>% 
  left_join(towns_recent_weeks %>% mutate(recent_rate = cases / (total_pop / 100000),
                                          recent_rate = ifelse(recent_rate < 0, 0, recent_rate)) %>% 
              select(town, recent_rate), by = "town") %>% 
  left_join(town_with_nursing %>% 
              filter(date == max(date)) %>% 
              select(town, nh_death_pct), by = "town") %>% 
  select(town, category, total_pop, `total cases`, `total deaths`,
         nh_death_pct, `town cases`, recent_rate,
         `town deaths`) %>% 
  mutate(nh_death_pct = formattable::percent(nh_death_pct, digits = 0),
         recent_rate = sprintf("%7.1f", recent_rate),
         `total cases` = scales::comma(`total cases`, accuracy = 1)) %>% 
  # use col.names to get column names in format_table
  select(town, category, `total population` = total_pop, `total cases`, 
         `recent cases per 100K` = recent_rate, 
         `total deaths`, `% deaths from nursing homes` = nh_death_pct, 
         `town cases`, `town deaths`) %>% 
  formattable::format_table(
    col.names <- c("x", "category", "total\npopulation", "total\ncases",
                   "recent cases\nper 100K", "total\ndeaths",
                   "care home deaths\nas % of total","trend in cases", "trend in deaths"),
     format = "html",
    align = c("l", "l", "r", "r", "r", "r", "r", "c", "c")
  ) %>%
  # kable_styling("striped", full_width = F) %>%
  # kable_styling(full_width = F) %>%
  # group_rows("group1", 1, 2) %>%
  # group_rows("group1", 1, 2) %>%
  # group_rows("group2", 3,3) %>%
  htmltools::HTML() %>%
  shiny::div() %>%
  sparkline::spk_add_deps()
# county_summary
town_summary
```
```{r county_towns, echo = FALSE}

for_range <- dph_towns %>% ungroup() %>% 
  filter(date == max(date, na.rm = TRUE)) %>% 
   mutate(`cases per 100K` = (cases / (total_pop / 100000)))
max_range = max(for_range$`cases per 100K`, na.rm = TRUE)
county_data <- town_geometries %>% 
              filter(county == params$county) %>% 
              left_join(dph_towns %>% 
                          filter(date == max(date, na.rm = TRUE)) %>% 
                          select(town, cases), by = "town") %>% 
              mutate(`cases per 100K` = cases / (total_pop / 100000))
```

The first map shows cases by town in `r params$county` County. The second map shows the percentage
of individuals below the poverty line.

```{r county_towns2, echo = FALSE, warning = FALSE}
p_cases_per100k <- ggplot(data = county_data, 
              aes(geometry = geometry)) +
  geom_sf() +
   geom_sf(aes(fill = `cases per 100K`, geometry = geometry)) +
  # scale_fill_gradient(low = "white", high = "grey") +
  # geom_sf_text(aes(label =town, color = "black", size = 3) +
  geom_sf_text(aes(label = town), size = 3, colour = "gray95") +
  coord_sf(datum = NA, label_axes = "----") +
  xlab("") + ylab("") + theme_minimal()  +
  scale_fill_viridis(direction = -1, limits = c(0, max_range), option = "viridis") +
  # theme(legend.position = "none") +
  labs(title = paste0(params$county, " County: Total Covid-19 Cases per 100K of Population by Town"),
       caption = "Sources: CT DPH, US Census, tidycensus package")

p_poverty <- ggplot(data = town_geometries %>% 
                      filter(county == params$county), 
              aes(geometry = geometry)) +
  geom_sf() +
   geom_sf(aes(fill = poverty_pct, geometry = geometry)) +
  # scale_fill_gradient(low = "white", high = "grey") +
  # geom_sf_text(aes(label =town, color = "black", size = 3) +
  geom_sf_text(aes(label = town), size = 3, colour = "gray95") +
  coord_sf(datum = NA, label_axes = "----") +
  xlab("") + ylab("") + theme_minimal()  +
  scale_fill_viridis(direction = -1, name = "poverty pct", 
                     limits = c(0, max(town_geometries$poverty_pct, na.rm = TRUE)), 
                     labels = scales::label_percent(accuracy = 1)) +
  # theme(legend.position = "none") +
  labs(title = paste0(params$county, " County: Percent Poverty by Town"),
       caption = "Sources: US Census, tidycensus package")
print(p_cases_per100k)
print(p_poverty)
```



